---
title: "Filtering and Mapping"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

*Author: Stuti Singh*

*Date: 2023-07-31*

*Description: This script performs R basic filtering and mapping using Leaflet and Geosphere*   



**Load the required packages**  
```{r}
rm(list=ls())
requiredPackages <- c('foreign','tidyverse', 'ggplot2', 'data.table', 'dplyr', 'tidyr', 'gmodels', 'Hmisc','arrow', 'sf', 'leaflet', 'htmlwidgets', 'gridExtra')
for(p in requiredPackages){
  if(!require(p, character.only = TRUE)) install.packages(p) 
}
```

**Read the files**   
*Here we are using a dummy dataset which contains establishments in Texas region providing information about employment, zipcode, location in the form of latitude and longitude and 30 other attributes.*

```{r}
dat <- read.csv("dummy_data.csv")
```

**Print first five rows to understand the data**  
```{r}
head(dat)
```


**Filter using a row value**    
*Here we are using a particular zipcode to filter all the rows corresponding to that zipcode value.* 
```{r}
filter_zip = dat %>%
  filter(dat$zipcode == 79118)
```


**Filter using name or a string in a row**    
*Here we are using the string "WAREHOUSE" to find a company which contains name warehouse.*

```{r}
substring_to_find <- "WAREHOUSE"
filter_by_name <- dat[grepl(substring_to_find, dat$company) , ]
  
# Display the filtered data
print(filter_by_name)
```

**Display the column names**
```{r}
colnames(dat)
```

**Filter using column names**
```{r}
Filter_columns <- dat[c("dunsnumber","company","address", "city" ,"state", "zipcode","hqzipcode","hqcity", "latitude","longitude","emp19", "empc19","emphere" ,"empherec", "industry", "NAICS19" )]

print(Filter_columns)
```


**Save the new dataset as a csv**
```{r}
write.csv(filter_by_name, file = "filter_by_name.csv", row.names = FALSE)
```


```{r}
dat$longitude <- paste0("-", dat$longitude)
dat$longitude <- as.numeric(dat$longitude)
head(dat)

```

**Creating a leaflet map using the latitude and longitude values**
```{r}
map_a <- leaflet() %>%
  addTiles() %>%
  addCircleMarkers(data = dat, fillColor = "green", fillOpacity = 0.6, weight = 0.5, group = ~zipcode)%>%
  addLayersControl(overlayGroups = unique(dat$zipcode), options = layersControlOptions(collapsed = FALSE))


map_a
```



**Filtering points within 100 miles from a given point**
```{r}
# Specify the coordinates of the center point (e.g., 40.0 latitude and -100.0 longitude)
center_lat <- 33.1078
center_lon <- -96.7033

# Specify the radius in miles
radius_miles <- 100

# Calculate the distance in miles between each marker and the center point
dat <- dat %>%
  mutate(
    distance_miles = geosphere::distGeo(matrix(c(center_lon, center_lat), ncol = 2), matrix(c(longitude, latitude), ncol = 2)) * 0.000621371
  )

# Filter the data to include only markers within the desired radius
data_within_radius <- dat %>%
  filter(distance_miles <= radius_miles)

# Create the Leaflet map
map_b <- leaflet(data = data_within_radius) %>%
  addTiles() %>%
  addCircleMarkers(lng = ~longitude, lat = ~latitude, radius = 2, color = "green", fill = TRUE, popup = ~company, group = ~zipcode)%>%
  addLayersControl(overlayGroups = unique(data_within_radius$zipcode), options = layersControlOptions(collapsed = FALSE))

map_b
```


**Save the map as html file**
```{r}
saveWidget(map_b, file = "MAP.html")

```